---
title: "Pratical Machine Learning Write Up"
author: "Sergio Regueira"
date: "26 de julio de 2015"
output: html_document
---

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

#Analysis
## Loading and cleaning Data

First of all, I load csv file in a variable

```{r}
library(caret)
library(parallel)
library(doParallel)
data <- read.csv("pml-training.csv",header=TRUE)
dataTest <- read.csv("pml-testing.csv", header = TRUE)
```

Clean up near zero variance features, columns with over 90% of NA and descriptive fields

```{r}
# REMOVE NA's
NAs <- apply(data,2,function(x) {sum(is.na(x))})
data <- data[,which(NAs < nrow(data) * 0.9)]

#REMOVE NZV
NZV <- nearZeroVar(data, saveMetrics = TRUE)
data <- data[, NZV$nzv == FALSE]

#REMOVE DESCRIPTIVE FIELDS (FIRST 7 COLUMNS)
data <- data[,7:ncol(data)]
```

Now, we convert Class column into factor.

```{r}
data$classe <- factor(data$classe)
```

## Splitting Data
To create the machine learning model we split the data in 60% for training and 40% for testing.

```{r}
trainIndex <- createDataPartition(data$classe, p=0.6, list = FALSE)
training <- data[trainIndex,]
testing <- data[-trainIndex,]
```

## Random Forest Model
The machine learning model generated is random forest (rf)

```{r}
set.seed(123)
registerDoParallel(makeCluster(detectCores()))
modelfit <- train(classe ~ ., method="rf",data=training)
```

## Accuracy
```{r}
rf_accuracy <- predict(modelfit, testing)
print(confusionMatrix(rf_accuracy,testing$classe))
```

## Cross Validation
```{r}
set.seed(123)
registerDoParallel(makeCluster(detectCores()))
control <- trainControl(method="repeatedcv",number=10, repeats=10)
modelCrossValidation <- train(classe ~ ., method="rf",data=training,trControl=control)
rf_CV_accuracy <- predict(modelCrossValidation, testing)
print(confusionMatrix(rf_CV_accuracy,testing$classe))
```

#PREDICT TEST CASES

```{r}
ptest <- predict(modelCrossValidation, dataTest)
answers <- as.vector(ptest)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}

pml_write_files(answers)
```
